---
title: "Spokane Crime Data"
output: html_document
---

```{r setup}

```

This walkthrough shows the steps needed to convert the Spokane Crime Data into a final dataset that is suitable for analysis. The R code is based on the `basecode.R` script by @kmigori. 


## Raw data files

The raw data is contained in two .zip archives - Crime.zip and Neighborhood.zip. The files can be downloaded manually using these links: 

[https://static.spokanecity.org/download/gistransfer/pubdata/Crime.zip](https://static.spokanecity.org/download/gistransfer/pubdata/Crime.zip)  
[https://static.spokanecity.org/download/gistransfer/pubdata/Neighborhood.zip](https://static.spokanecity.org/download/gistransfer/pubdata/Neighborhood.zip)  

Alternatively, you can enter the following into the R console. 

```{r, echo = TRUE, eval = FALSE}
library(httr)  
crime_url <- "https://static.spokanecity.org/download/gistransfer/pubdata/Crime.zip"  
neighborhood_url <- "https://static.spokanecity.org/download/gistransfer/pubdata/Neighborhood.zip"  

GET(crime_url, write_disk("Data/Crime.zip", overwrite=FALSE))  
GET(neighborhood_url, write_disk("Data/Neighborhood.zip", overwrite=TRUE))  


```  

Either way, the .zip files should be saved in the /Data folder in the repository's root directory.


## Inspect the .zip archive

The resulting .zip archives can be inspected and unzipped manually, or at the R console using this code. 

```{r, echo = TRUE, eval = FALSE}
unzip("Data/Crime.zip", list = TRUE)
unzip("Data/Neighborhood.zip", list = TRUE)

unzip("Data/Crime.zip", exdir = "Data/Crime")
unzip("Data/Neighborhood.zip", exdir = "Data/Neighborhood")

```


```{r inspect zip, echo = FALSE, eval = TRUE, comment = ""}
unzip("../Data/Crime.zip", list = TRUE)

unzip("../Data/Neighborhood.zip", list = TRUE)

#unzip"../Data/Crime.zip", exdir = "Data/Crime")
#unzip"../Data/Neighborhood.zip", exdir = "Data/Neighborhood")
```


## Load the data

This is where Kristian's code comes into play. It has been modified slightly to account for the data files being saved in the /Data directory. 

```{r load shape, echo = TRUE, eval = FALSE}
#this is an R script to import Spokane City crime data, merge it with Neighborhoods, and export it as a flat file
rm(list=ls())
library(rgdal)
library(maptools)

#reading in Crime shapefile
Crime=readOGR(dsn="Data/Crime",layer="Crime")
#reading in Neighborhood shapefile
Neighborhood=readOGR(dsn="Data/Neighborhood",layer="Neighborhood")

#Extracting Neighborhood designation for each point by location
Crime_byNeighborhood=over(Crime,Neighborhood)
#appending rest of data to each point
Crime_byNeighborhood=spCbind(Crime,Crime_byNeighborhood)
#saving resulting dataset as csv
write.csv(file="Data/Spokane_CrimeData_Neighborhood.csv",Crime_byNeighborhood)
```

